---
title: "Choosing a home purchase location and high school for optimal University of California admissions"
client:
author: "Quynh Tran"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> "Education is the passport to the future, for tomorrow belongs to those who prepare for it today."
([via](http://www.blackpast.org/1964-malcolm-x-s-speech-founding-rally-organization-afro-american-unity))



http://admitguide.com/uc-acceptance-rate-by-high-school/

As you would expect, admissions to the UCs vary considerably by high school. I was shocked to learn that our local high school, Castro Valley High had a much lower acceptance rate to UC Berkeley, for example, than most other public high schools in Alameda County. Despite its overall historical relatively high API score, Castro Valley high school students could not get into college, relatively. ts district?

This answer could use data from 1994 to today, include race because the acceptance rates differ across race, and the different UC campuses. Merced is the easiest to gain admissions, while UCLA is the most difficult.

One challenge for me is that I am used to individual level data and predictive modeling and have not worked with summary data in any predictive modeling. However, the statistics involved with this project is very simple and I do not expect any modeling. The data wrangling and coding will be more challenging because of the way the data is released and that I am still learning R.


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Stake holders
This proposal is for (fictional) real estate firm River Trends that specializes in home purchases for foreign clients. The majority of the clients are interested in real estate in areas with strong  


## Purpose

One of the first questions asked by people moving to the Bay Area is where they should live and buy a home. Invariably the answer comes down to home affordability and school district because good schools is the driving force for real estate values.

The purpose of this project is to identify which locations in Alameda County to purchase a home and its corresponding public high school that would achieve the optimal chances for admissions to the University of California campuses.

## Benefits

2018 marks the 150th anniversary of the creation of the University of California system. At the same time, applications to the University of California campuses are at an all time peak with an increase of 5.7 percent system-wide. Five UC campuses -- Berkeley, Irvine, Los Angeles, San Diego and Santa Barbara -- each received more than 100,000 applications this past year. San Diego's total is up 9.7 percent. Davis received some 95,000 applications, up 8.6 percent. UC Riverside saw the largest percentage increase -- 12.2 percent. One of the most competitive campuses, Berkeley, only received an increase of 4.6 percent. (By way of comparison, Harvard University received just under 40,000 applications last year.) Berkeley and UCLA are the most competitive campuses, accepting about one in six applicants. ^[https://www.insidehighered.com/admissions/article/2018/03/26/reports-circulate-even-more-difficult-year-be-admitted-leading]

Admission offers, however, depended on where the applicant attended high school. Selecting a high school with higher acceptance rate to the target University of California campus is a prudent choice for families in their selection of home purchase. For example, the acceptance rate to Berkeley for Castro Valley High School applicants is about 11 percent, whereas applicants from Hayward High School 

Parents are increasingly stressed over UC admissions but few notice the discrepancy in admissions rate between choice of high school. ^[https://www.sfchronicle.com/opinion/article/UC-needs-to-be-transparent-about-who-gets-in-12757352.php]

## Success criteria

## Estimated risks

College admissions rate is rarely a significant variable when families consider their home purchase. Many, however, do seek the general quality of the local school district and elementary/ middle schools. Yet ignoring the quality of the public high school may be the most significant oversight when making a home purchase decision. The high school education builds upon the curriculum of the elementary and middle schools. It is the culmination of the entire public education experience, which makes it the strongest marker of the quality of the school district.  

By identifying the most desireable public high schools based on target UC campus admissions, 

## Audience or users

The audience for this project are potential homebuyers with children in public high school, heading into public high school, or those thinking long term about public high schools and admissions into the University of California campuses.

Users of this information also include real estate firms who can steer home purchasers towards the locations and their respective high schools for optimal admissions into their most desired University of California campus.

Previously the California Department of Education posted an API or Academic Performance Index score for each public school and school distrcit in the state. The composite test score had a maximum of 1000 with 800 deemed as a good goal. I cross-referenced the test scores with housing prices for the area. I quickly discovered that Castro Valley offered the best test score to housing price ratio. Castro Valley had schools whose API scores rivaled Palo Alto, but for up to $300,000 less. We bought a home and moved to Castro Valley.

Twelve years later, I find that people are still asking the same question but they are also thinking forward to college admissions. School districts no longer offer an API score for comparison between schools and districts. However, the University of California system recently started to release its admissions data. They are only summary statistics and not individual data. But the most interesting information is the names of each high school and how many students apply and are admitted to each UC campus going back to 1994. The summary statistics also include the race variable.

## Timeline

The timeline for this project is estimated at 8 weeks. 

 Twelve years ago, I faced this question when pregnant with my first child. We were living in Mountain View, where Google was beginning to take hold of the town. It was at the peak of the housing market in 2006. The house near our apartment, at 800 sq. ft. was selling for $800,000. 


Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

math expressions, e.g. $Y = X\beta + \epsilon$, 

`knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```


